---
title: "Olympics Games"
author: "Wenjia Xie"
date: "December 18, 2018"
output: ioslides_presentation
---


```{r setup, include=FALSE}
pacman::p_load(
  "tidyverse",
  "data.table",
  "benford.analysis",
  "gapminder",
  "gapminder",
  "plotly"
)
```

## The introduction of the dataset
This project used a historical dataset on the modern Olympic Games, including all the Games from Athens 1896 to Rio 2016.

The Olympic data is the result of an incredible amount of research by a group of Olympic history enthusiasts and self-proclaimed 'statistorians'. Randi H Griffin scraped this data from the webpage and consolidated their work into a convenient format for data analysis.

Please check out their blog for more information:
![www.sports-reference.com]

```{r,include=FALSE}
athlete<-read_csv("athlete_events.csv")
region<-read_csv("noc_regions.csv")
athlete[which(athlete$Sex=="M"),3]<-"Male"
athlete[which(athlete$Sex=="F"),3]<-"Female"
female<-athlete %>% 
  select(Sex,Height,Weight,Games,Sport) %>% 
  filter(Sex=="F") %>% 
  na.omit()
male<-athlete %>% 
  select(Sex,Height,Weight,Games,Sport) %>% 
  filter(Sex=="M") %>% 
  na.omit()
```


## sport
```{r,echo=FALSE}
number<-athlete %>%
  select(Year,Name,Sex,Height,Weight,Season,Sport) %>%
  filter(Season=="Summer") %>%
  filter(Sex=="Male") %>%
  group_by(Sport) %>%
  summarise(number=n())


sport<-athlete %>%
  select(Year,Name,Sex,Height,Weight,Season,Sport) %>%
  filter(Season=="Summer") %>%
  filter(Sex=="Male") %>%
  unique(by=c(Year,Name)) %>%
  na.omit() %>%
  group_by(Sport) %>%
  summarise(mweight=mean(Weight),mheight=mean(Height)) %>%
  left_join(number,by="Sport")

sport$weight<-round(sport$mweight,2)
sport$height<-round(sport$mheight,2)

p <- plot_ly(sport,
             x = ~mweight,
             y = ~mheight,
             color = ~Sport,
             colors='Paired',
             size = ~number,
             type = 'scatter',
             mode = 'markers',
        marker = list(symbol = 'circle', sizemode = 'diameter',
                      line = list(width = 2, color = '#FFFFFF')),
        text = ~paste('Sport:', Sport, '<br> Mean Height:',
                      round(mheight,2),'<br>Mean Weight:', round(mweight,2),
                      '<br> Number of Athletes:', number)) %>%
  layout(title = 'Mean Weight vs. Mean Height,1986-2016',
         xaxis = list(title = 'Mean Weight',
                      gridcolor = 'rgb(255, 255, 255)',
                      range = c(60,95),
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwidth = 2),
         yaxis = list(title = 'Mean Height',
                      gridcolor = 'rgb(255, 255, 255)',
                      range = c(160,200),
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwith = 2),
         paper_bgcolor = 'rgb(243, 243, 243)',
         plot_bgcolor = 'rgb(243, 243, 243)') %>%
         layout(showlegend = FALSE)

p


```

## country
```{r,echo=FALSE}
Gcountry<-athlete %>%
      select(Team,NOC,Year,Season,Sport,Games,Event,Medal) %>%
      filter(Season=="Summer") %>%
      filter(Medal=="Gold") %>%
      unique(by=c("NOC","Year","Event")) %>%
      group_by(Year,NOC) %>%
      summarise(nmedal=n())




l <- list(color = toRGB("grey"), width = 0.9)  ### light grey boundaries


g <- list(
      showframe = FALSE,
      showcoastlines = FALSE,
      projection = list(type = 'Mercator')
    )     ### specify map projection/options

p <- plot_geo(Gcountry) %>%
      add_trace(
        z = ~nmedal,
        color = ~nmedal,
        colors = 'Blues',
        frame = ~Year,
        locations = ~NOC,
        marker = list(line = l)
      ) %>%
      colorbar(title = paste("Number of gold Medal")) %>%
      layout(title ='Evolution of the Olympics over time',
             geo = g)

p

```



## Benford Analysis
```{r,echo=FALSE}
height<-athlete %>% 
  select(Name,Sex,Height) %>%
  unique() %>% 
  na.omit()

fheight<-height %>% 
  filter(Sex=="Female")



bfd.gdp <- benford(fheight$Height,1)
plot(bfd.gdp)




```



# The distrbution of Women's height
```{r,echo=FALSE}
p<-ggplot(height,aes(x=Height,color=Sex))+
  geom_density(size=1.2,adjust=1.5)+
  geom_histogram(aes(y=..density..,fill=Sex), alpha=0.1,
                position="identity",bins=90)+
  theme_minimal()
ggplotly(p)

fheight<-height %>%
  filter(Sex=="Female")
```


# the QQplot of Women's height
```{r,echo=FALSE}
qqnorm(fheight$Height);qqline(fheight$Height, col = 2)

```
